\section{Legal issues and GDPR}
\textbf{Privacy by Design (PbD)} is an approach to system engineering which promotes privacy through out the whole engineering process. In particular, it takes human values into account in a well-defined manner through the whole process, it's not about data protection but designing so data doesn't need protection, and the root principle is based on enabling service without data control transfer from the citizen to the system. The PbD main principles are the following :
\begin{itemize}
\item \textbf{Proactive} : it means that we are taking into account privacy from the early beginning of the design phase.
\item \textbf{Privacy as the default setting} : it means that privacy is taken into account in all the stages of the design.
\item \textbf{Privacy embedded into design} : the privacy is directly interpreted as a requirement in the specification of the system.
\item \textbf{Full functionality} : it means that we are improving the quality of our system and not diminishing it (this is also called Positive-sum). So if we increase privacy we decrease security and vice versa (this is also called Zero-sum).
\item \textbf{End-to-End security} : it's a way for explicitly stating that we are considering the standard privacy principles of security.
\item \textbf{Visibility and transparency} : it means that we reveal whatever it's necessary to the users.
\item \textbf{Respect for user privacy} : it means that most the information that are collected is done focusing on the users (User centricity).
\end{itemize}
According to PbD supporters data minimization is the most important safeguard in protecting personally identifiable information and the use of cryptography, de-identification techniques and data aggregation are absolutely critical. PbD has been critiqued as "vague" and leaving "many open questions" about their application when engineering systems. It has also been pointed out that PbD is similar to voluntary compliance. The major limitations of this approach are :
\begin{itemize}
\item Current definitions of privacy by design do not address the methodological aspect of system engineering that we should use to implement that.
\item The concept also does not focus on the role of the actual data holder, but on that of the system designer. This role is not known in privacy law, so the concept of Privacy by Design is not based in law. This in turn undermines the trust by data subjects, data holders and policy makers.
\end{itemize}
The European Commission in January $2012$ presented a proposal to ensure a coherent framework and a harmonized system in EU matters. It consists of two different tools :
\begin{itemize}
\item a proposal for a regulation concerning "the protection of individuals with regard to the processing of personal data and the free movement of such data", aimed at regulating the processing of personal data in both the private and public sector.
\item a proposal for a directive addressed to the regulation of prevention, conflict and repression of crimes, and to the enforcement of criminal penalties.
\end{itemize}
The legislative process for such proposal has been concluded in $2016$ and it will be enforced starting from $2018$. This proposal goes under the name of \textbf{General Data Protection Regulations (GDPR)}. In the GDPR we can distinguish between two roles : \textbf{data controller}, which is the person or body that, alone or jointly with others, determines the purpose and means of the processing of personal data, and \textbf{data processor}, which is a legal person or body which processes personal data on behalf of the controller. In particular, the data processing must comply with the $6$ general GDPR principles :
\begin{itemize}
\item \textbf{Lawfulness, fairness and transparency}
\item \textbf{Purpose limitation} : it means that personal data must be collected for specified, explicit and legitimate purposes and not further processed in a manner that is incompatible with those purposes.
\item \textbf{Retention} : it means that personal data must be kept in an identifiable format only for as long as necessary. We need to be clear about the length of time data will be kept and reason for same.
\item \textbf{Integrity and confidentiality} : the personal data must be kept secure, in such a way that it doesn't permit unauthorized access, intentionally or accidentally.
\item \textbf{Data minimization} : the personal data must be adequate, relevant and limited to achieve the purpose.
\item \textbf{Accuracy} : the personal data must be accurate and up to date. The longer personal data is held, the more likely it will be inaccurate/out-of-date. For this reason, staff must ensure that local procedures are in place to ensure high levels of data accuracy, including periodic review and audit.
\end{itemize}
An important aspect is that if the data is anonymized, then GDPR doesn't apply. Instead, if the data is pseudonymized then GDPR does apply. GDPR requires businesses to implement "technical and organizational measures to provide appropriate protection to the personal data they hold". This requires to secure the Personal Identifiable Information (PII) and Personal Health Information (PHI) to prevent unauthorized access such that in case of unauthorized access the data they get is unintelligible. To reach this goal typically companies uses several measures such as encryption in order to guarantee an appropriate level of security. In particular, GDPR expressly states that such measures include :
\begin{itemize}
\item The \textbf{pseudonymization} and \textbf{encryption} of personal data.
\item Measures to ensure resilience of systems and services processing data.
\item Measures that allows businesses to restore the availability and access to the data in the event of a breach.
\item Frequent testing of the effectiveness of the security measures.
\end{itemize}
When encrypting personal data, in accordance with Article $4$ No. $5$ GDPR, the encryption key is the "additional information" which is "kept separately" and "subject to technical and organizational measures". Thus the safety measures must ensure that the personal data are not attributed to an identifiable natural person. GDPR includes a definition of "pseudonymization". According to Article $4$ No. $5$ GDPR, pseudonymization means the processing of personal data in such a way that the personal data can no longer be attributed to a specific data subject without the use of additional information, provided that such additional information is kept separately and is subject to technical and organizational measure to ensure that the personal data are not attributed to an identifiable natural person. Indeed, pseudonymization shall, like encryption, be one of the "appropriate safeguards" of Article $6$ Par. $4 (e)$ GDPR. The pseudonymization main limitations are the following : implements methodologies to guarantee pseudonymization might lead to a huge cost, since to manage them is typically required a qualified staff, and there not exist a default method to guarantee pseudonymization. The GDPR main key implications are the following :
\begin{itemize}
\item \textbf{Increased fines} : if one or more rules defines by GDPR are not satisfied by a company, then the regulators can imposes fines of up to $4 \%$ of the company annual turnover. In this case, regulators may perform audits, issue warnings or a temporary ban on processing.
\item \textbf{Proof of compliance} : it means that companies must demonstrate they are compliant by : evidencing that they comply the $6$ GDPR principles and processing conditions,documenting suitable policies that set out how they process personal data, performing Privacy Impact Assessments and implementing technical security measures.
\item \textbf{New rights} : this section contains several rights such as right to be forgotten, to data portability, to access and rectify personal data within $30$ days, etc.
\item \textbf{Privacy by Design, Privacy by Default} : it's mandatory to implement Pivacy by Design to ensure privacy and data protection is a key consideration during the entire life cycle of any project. Privacy by Default means to consider privacy as the default setting and directly embedded into the design.
\item \textbf{Data Protection Officers (DPO)} : it's the authority that send reports to highest levels of managements and it's mandatory in certain cases.
\item \textbf{Privacy Impact Assessments (PIA)} : It's mandatory for high risk personal data processing and in some cases consulting the supervisory authority is required.
\item \textbf{Privacy notices} : It provides an increase of mandatory amount of information included in privacy notices and it's supplied to the individual at the time they provide personal data. If processing is for a new purpose then a prior notification must be given. It must be concise, transparent, intelligible and easily accessible.
\item \textbf{Consent} : Consent must be freely given, specific, informed and unambiguous. It may be withdrawn at any time and must be explicit for sensitive personal data or for data transfer outside the EU.
\item \textbf{Mandatory breach notifications} : whenever a data breach occurs then they must notify the interested users within a specified time window.
\item \textbf{Obligations for data processors} : they are new obligations for data processors in order to be more responsible and liable.
\item \textbf{Extra-territorial scope} : the GDPR applies to data controllers and processors established in the EU and organizations that target EU citizens.
\end{itemize}
A \textbf{Data Protection Impact Assessment (DPIA)} is required any time we begin a new project that is likely to involve "a high risk" to other people's personal information. For instance, when we use large-scale processing operations which aim to process a considerable amount of personal data and which are likely to result in a high risk for privacy. A DPIA should be done when a data breach occurs and the organizations must show that they did their best to avoid the data breach and also because fines depends on the DPIA evaluation. The problem is that the GDPR doesn't define a particular process for a DPIA, and thus we need some requirements to define the DPIA. For this reason, it allows for flexibility in terms of scope and scale (the organizations can develop their own internal approach) and allows for alignment with existing project management approaches. The typical approach for defining a DPIA is through a systematic description of what could be the possible dangers (i.e. possibility for an internal or external intruder for accessing data), an evaluation of the necessity and proportionality of the processing operations in relation to the purposes, an evaluation of the risks to the rights and freedoms of data subject referred estimating likelihood and severity, measures envisaged to address risks including mechanisms to ensure protection of personal data and mechanisms to demonstrate compliance with the GDPR and the controller shall seek the views of data subjects or their representatives on the intended processing.\\\\The GDPR defines the following characteristics of the regulations for privacy for individuals :
\begin{itemize}
\item \textbf{easier access to your own data} : it means that individuals will have more information on how their data is processed and this information should be available in a clear and understandable way.
\item \textbf{a right to data portability} : it means that will be easier to transfer your personal data between service providers.
\item \textbf{a "right to be forgotten"} : when you no longer want your data to be processed, and provided that there are no legitimate grounds for retaining it, the data will be deleted.
\item \textbf{the right to know when your data has been hacked} : it means that companies and organizations must notify the national supervisory authority of serious data breaches as soon as possible so that users can take appropriate measures.
\item \textbf{Data protection by design and by default} : the first one requires that data protection is designed into the development of business processes for products and services. Whereas the second one means that the default settings should be those that provide the most privacy.
\item \textbf{Stronger enforcement of the rules} : it means that data protection authorities will be able to fine companies who do not comply with EU rules up to $4 \%$ of their global annual turnover.
\end{itemize}
While the characteristics of the regulations for privacy for business defined by the GDPR are the following :
\begin{itemize}
\item \textbf{One continent, one law} : the regulation will establish one single set of rules which will make easier and cheaper for companies to do business in the EU.
\item \textbf{One-stop-shop} : business will only have to deal with one single supervisory authority. It will favor the cooperation between the data protection authorities on issues of interest for all of Europe. In this way, business will make profit from faster decisions, from one single interlocutor and from less red tape.
\item \textbf{European rules on European soil} : the companies outside of Europe will have to apply the same rules when offering services in EU.
\item \textbf{Risk-based approach} : the rules will avoid a "one-size-fits-all" obligation and rather tailor them to the respective risks.
\item \textbf{Rules fit for innovation} : data protection by design will guarantee that data protection safeguards are built into products and services from the earliest stage of development. Furthermore, privacy-friendly techniques such as pseudonymization, anonymization and encryption will be encouraged. This will encourage the use of big data analytics. The fundamental aspect is that when the personal data is collected for one or more purposes it shouldn't be further processed in a way that is incompatible with the original purposes. Deciding whether or not is fair a new purpose incompatible with the original purposes is another fundamental factor. In particular, the fairness will consider factors such as the effect on the privacy individuals and whether an individual has a reasonable expectation that their personal data will be used in the new way.
\end{itemize}

\subsection{Data protection officer}
The data protection officer is a designed person within an organization that collects the personal data of Union citizens, who is responsible for making sure that the organization follows the new regulations. The DPO main tasks are the following :
\begin{itemize}
\item \textbf{inform and advise} the owner of the treatment about their obligations under the European regulations.
\item \textbf{check the implementation and application of the regulations} and provide opinions on the assessment of the impact on data protection.
\item acts as a contact point about any issue related to the processing of their data or the exercise of their rights.
\item act as a contact point for the Authority for the protection of personal data or, where appropriate, refer to the Guarantor at its own initiative.
\end{itemize}
However, the DPO in order to perform the previous tasks needs to have the following requirements :
\begin{itemize}
\item Adequate knowledge of legislation and personal data management practices.
\item Carry out its functions in full independence and in the absence of conflicts of interest.
\item Operate in the employ of the owner or responsible on the basis of a contract. The owner or manager of the treatment will have to make available to the head of personal data protection (DPO) human and financial resources necessary to perform its tasks.
\end{itemize}
A DPO must be compulsorily appointed by government departments and agencies (with the exception of judicial authorities), all person whose main business consists of treatments which require regular and systematic monitoring of those concerned, and all persons whose main business is processing sensitive data concerning health or sex life, genetic, judicial and biometrics.

\subsection{Italy regulations}
In Italy with the legislative decree of June $30$th $2003$ established the regulations and laws for protecting sensitive data according to the fast spreading of application technologies in the world. In particular, the data considered sensible by the regulations are the following :
\begin{itemize}
\item \textbf{Sensitive data} : personal data revealing ethic origin, religious beliefs, philosophical or other beliefs, political opinions, health and sex life, etc.
\item \textbf{Judicial data} : personal data relating to criminal records, the register of offense-related administrative sanctions and the relevant current charges, or as an accused or suspected person of the criminal procedure code.
\end{itemize}
Such data needs to be processed according to the following rules :
\begin{itemize}
\item Anyone has the right to protection of personal data concerning him.
\item The processing of personal data will be respect for human rights and fundamental freedoms and dignity, with particular reference to privacy and the right to identity the protection of personal data.
\item The information systems and programs are configured to minimize the use of personal data and identification data, in order to rule out their processing if the purposes required in the individual cases can be achieved by using either anonymous data or mechanisms permitting the identification of the individual only in case of need.
\end{itemize}
The rights of individuals data are the following :
\begin{itemize}
\item \textbf{The right of access} to personal data, including the right to obtain confirmation of the existence of data concerning him, the right to have their communication in intelligible form.
\item \textbf{The updating, correction or integration of data}.
\item \textbf{The cancellation, anonymization or blocking of data} processed in violation of the law.
\item \textbf{The right to oppose} : for legitimate and documented reasons to the treatment, even if carried out in a legitimate way or if the processing carried out for the purpose of sending advertising materials or direct selling or for carrying out market research. 
\end{itemize}
In Italy there is a collegian body composed of four members elected by Parliament called \textbf{Authority (Garante)}. Such entity must communicate data processing and preliminary consent of specific actions such as the usage of detection systems of biometric, data processed with the help of electronic instruments to define the profile or personality, etc. Whereas the obligations for individuals are disclosure of information, authorization and consensus of personal data treatment, external information policy (this could a problem in processing information). The major problems for companies and administrations are that they have to follow a well defined and strict set of rules regarding non disclosure of information and must have an explicit consent from the users in order to process the data for a specific and explicit purpose. We understand that such process is very long and difficult in order to be fully compliant with the Italian laws. Other complications may come from \textbf{Bring your own device (BYOD)} approach, where companies monitors their employees together with their devices for security reasons, but this must always be done in accordance with the law. Such monitoring should be restricted to use of the device only within the employment context, because doing that outside is disproportional.